Cartograms with Dynamic Features

### 2022-09-15 paper/section/data.tex
fix: typos

\bobgraph{Clinical Commissioning Groups Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs changes over time due to the NHS re-organization. The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations took place in 2021 and 2022.

\bobgraph{River Features:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Great Ouse in England. These rivers were chosen as they are well-known rivers and pass through regions with dense population, and provide informative geographical and topological cues.

@@

After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections, and convert them into GeoJSON files. Finally, mapshaper \cite{blochMapshaper} is used to merge adn convert the GeoJSON files into a TopoJSON \cite{TopoJSON} file. TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation. See \autoref{table:pre-processing_result} for the pre-processing result.

>>>>

\bobgraph{Clinical Commissioning Groups Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs changes over time due to the NHS re-organization. The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.

\bobgraph{River Features:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent, and River Great Ouse in England. These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues.

@@

After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections, and convert them into GeoJSON files. Finally, mapshaper \cite{blochMapshaper} is used to merge and convert the GeoJSON files into a TopoJSON \cite{TopoJSON} file. TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation. See \autoref{table:pre-processing_result} for the pre-processing result.

### 2022-09-15 paper/section/evaluation.tex
fix: typos

\begin{itemize}
    \item Under 75 mortality from cardiovascular disease
    \item Emergency admissions for alcohol related liver disease
\end{itemize}

@@

\textbf{T1:} The participants were asked to familiarize with the design, including the user options and the interaction methods.
\textbf{T2:} The participants

>>>>

\begin{itemize}
    \item Under 75 mortality from cardiovascular disease
    \item Emergency admissions for alcohol-related liver disease
\end{itemize}

@@

\textbf{T1:} The participants were asked to familiarize themselves with the design, including the user options and the interaction methods.
\textbf{T2:} The participants

### 2022-09-15 paper/section/introduction.tex
fix: typos

Cartograms are representations of hybrid geographical and abstract data based on a value-by-area mapping technique combining statistical and geographical information \cite{dent2009Cartography}. Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision making \cite{gao2020Visualising, sack2021Visualizing}.

Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \autoref{table:accuracy}). For this project we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space. Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy. We implement a new cartographic layout algorithm that includes these topological features into the layout of the nodes representing geographical regions. To minimize geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout. We then apply the algorithm to a real world case-study using an Electronic Health Records (EHR) dataset to evaluate of our algorithm. We also present a user study demonstrating its effectiveness.

@@

One of the major challenges involved is how to develop a layout algorithm that handles different shapes. In other words, the layout algorithm is novel because it handles different types of nodes -- rectangular representing regions and polylines representing rivers. Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by dynamic topological features while minimizing the geographical error.

>>>>

Cartograms are representations of hybrid geographical and abstract data based on a value-by-area mapping technique combining statistical and geographical information \cite{dent2009Cartography}. Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision-making \cite{gao2020Visualising, sack2021Visualizing}.

Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \autoref{table:accuracy}). For this project, we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space. Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy. We implement a new cartographic layout algorithm that includes these topological features into the layout of the nodes representing geographical regions. To minimize geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout. We then apply the algorithm to a real-world case study using an Electronic Health Records (EHR) dataset to evaluate our algorithm. We also present a user study demonstrating its effectiveness.

@@

One of the major challenges involved is how to develop a layout algorithm that handles different shapes. In other words, the layout algorithm is novel because it handles different types of nodes -- rectangular representing regions and polylines representing rivers. Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by dynamic topological features while minimizing geographical error.

### 2022-09-15 paper/section/related.tex
fix: typos

\thumbnail{rectangular}{
	Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}. Depending on the variant, a rectangular cartogram may trade off between statistical and topological accuracy.
}

@@

\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States. \citea{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms. \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China. \citea{tong2018cartograms} use a Demers cartogram to visualize health-related data by CCG regions in England, the work introduces a novel technique to remove overlap of squares based on topological features, aiming to improve both geographical and topological accuracy. \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US. See \autoref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.

Our work extends the algorithm described by \citea{tong2018cartograms} which incorporates a static topological feature into Demers cartograms. Our work enhances that of \citeauthor{tong2018cartograms} in three ways. We introduce multiple features (rivers) into the layout (as opposed to a single), making topological features dynamic further improves the legibility and geographical accuracy. We improve the algorithm for resolving stalemates.  The way we evaluate the cartograms is also different. \citeauthor{tong2018cartograms} count river crossings to evaluate error. Here our focus is on readability, thus we include a preliminary user study.

@@

One of the disadvantages of Dorling and Demers cartograms is legibility. The layout algorithms may displace regions far from their original position and making the maps more difficult to interpret. Our work extends Demers cartograms. We introduce a new type of dynamic topological feature, a river, aiming to improve the interpretation, readability, and geographical accuracy of this class of cartograms.

>>>>

\thumbnail{rectangular}{
	Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}. Depending on the variant, a rectangular cartogram may trade-off between statistical and topological accuracy.
}

@@

\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States. \citea{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms. \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China. \citea{tong2018cartograms} use a Demers cartogram to visualize health-related data by CCG regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy. \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US. See \autoref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.

Our work extends the algorithm described by \citea{tong2018cartograms} which incorporates a static topological feature into Demers cartograms. Our work enhances that of \citeauthor{tong2018cartograms} in three ways. We introduce multiple features (rivers) into the layout (as opposed to a single), making topological features dynamic and further improving the legibility and geographical accuracy. We improve the algorithm for resolving stalemates.  The way we evaluate the cartograms is also different. \citeauthor{tong2018cartograms} count river crossings to evaluate error. Here our focus is on readability, thus we include a preliminary user study.

@@

One of the disadvantages of Dorling and Demers cartograms is legibility. The layout algorithms may displace regions far from their original position and make the maps more difficult to interpret. Our work extends Demers cartograms. We introduce a new type of dynamic topological feature, a river, aiming to improve the interpretation, readability, and geographical accuracy of this class of cartograms.
